Raw Data Comparison

### 2019-09-11 Raw Data Comparison.Rmd
...

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(BarrettsProgressionRisk)
library(reshape2)
library(gridExtra)
library(e1071)

@@

dim(vfit.data)
dim(tfit.data)
```

@@

wilcox.test(t.fit.disp$mean, v.fit.disp$mean)

@@

wilcox.test(t.raw.disp$mean, v.raw.disp$mean)

@@

# First processing steps
```{r}
if (file.exists('prepped.Rdata')) {
  load('prepped.Rdata')
} else {
  t.prepped = BarrettsProgressionRisk:::.prepRawSWGS(traw.data, tfit.data, blacklist, verbose=T, plot=F)
  dim(t.prepped$data)

@@

all.val %>% dplyr::filter(Samplename == samples$Sample[2])
```

@@

@@

Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
```{r, fig.width=10}
per.pt.tile<-function(df) {
  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('mean' = 'V1')
  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('sd' = 'V1')
  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('var' = 'V1')
  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('IQR' = 'V1')

@@

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)

library(tidyverse)
library(BarrettsProgressionRisk)
library(reshape2)
library(gridExtra)
library(kableExtra)
library(pander)
library(e1071)

@@

##dim(vfit.data)
##dim(tfit.data)
```

@@

pander(wilcox.test(t.fit.disp$mean, v.fit.disp$mean), caption='Per pt mean comparison (fitted values)')

# Within the cohorts they're comparable
x = t.fit.disp$mean[sample(500,replace = F)]
pander(wilcox.test(x[1:250], x[251:500]),caption='Within training (sampled)')

y = v.fit.disp$mean[sample(100,replace = F)]
pander(wilcox.test(y[1:50], y[51:50]), caption='Within validation (sampled')

pander(wilcox.test(t.fit.disp$var, v.fit.disp$var), catpion='Per pt var comparison (fitted values)')

@@

pander(wilcox.test(t.raw.disp))

@@

The samples from the training cohort that have a mean>300 (outliers) are progressors, but did not all predict high risk in LOO.

# First processing steps

In the initial processing the raw counts are adjusted using the fitted values.

```{r}
if (file.exists('prepped.Rdata')) {
  load('prepped.Rdata')
} else {
  t.prepped = BarrettsProgressionRisk:::.prepRawSWGS(traw.data, tfit.data, blacklist, verbose=T, plot=F)
  dim(t.prepped$data)

@@

all.val %>% dplyr::filter(Samplename == samples$Sample[2])

rm(prr500, prr250, pd500, pd250, seg500, seg250)
```

@@

5MB bins.  For most bins this involves a weight average of 2 segments that overlap the bin.

@@

## Per patient

Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
```{r, fig.width=10}

per.pt.tile<-function(df) {
  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('mean' = 'V1')
  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('sd' = 'V1')
  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('var' = 'V1')
  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('IQR' = 'V1')

@@

t.per.pt.scale = as_tibble(t(apply(as.matrix(training.tiles[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=training.tiles$Sample, .before=1)
colnames(t.per.pt.scale) = colnames(training.tiles)

v.per.pt.scale = as_tibble(t(apply(as.matrix(val.tiles[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=val.tiles$Sample, .before=1)
colnames(v.per.pt.scale) = colnames(val.tiles)

t.pt = per.pt.tile(t.per.pt.scale) %>% mutate(cohort='train')
v.pt = per.pt.tile(v.per.pt.scale) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.pt, v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + ylim(-.5,.5) +
    geom_jitter(width=0.1) + geom_boxplot(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean per patient'),
  ggplot(bind_rows(t.pt, v.pt), aes(x=cohort, y=sd, group=cohort, fill=cohort)) + 
    geom_point() + geom_boxplot(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='StdDev per patient'),
  ncol=2, top='Per patient tiled values, after scaling per patient')

@@

## Per bin

@@

### Per bin, after scaling per-pt

t.dist = tile.dist.measure(t.per.pt.scale) %>% mutate(cohort='train')
v.dist = tile.dist.measure(v.per.pt.scale) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'), 
  ncol=2, top='Per bin tiled values on per-pt scaled')
